### 2016-08-02 ClimaTrans/climatrans.Rmd

```{r,echo=TRUE}
library(esd)
library(ncdf)
## Information about the system and session
```

@@

### **Daily station data**
```{r,echo=TRUE}
## Load precipitation data extracted from the gridded IMD data set:
load('climatrans.pr.rda')
## Display the structure of the data as
str(climatrans.pr)
```
##**Sensitivity tests**

@@

```{r}
wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
season<-5:10
fw <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetfreq')
mu <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetmean')
nd <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'nv')
q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wq95')
```

@@

```{r, fig.width=6, fig.height=6}
## Wet-day mean
Pr <- PrexpPr(mu,fw,x0=50) 
plot(Pr,new=FALSE,errorbar=FALSE)
for(i in 1:3) lines(trend(subset(Pr,is=i)))
grid()
```

@@

```{r, fig.width=6, fig.height=6}
## Aggregate to mean seasonal cycle and annual mean values
pt <- annual(climatrans.pr,'sum')
attr(pt,'unit') <- 'mm/season'
plot(pt,new=FALSE)
for(i in 1:3) lines(trend(subset(annual(climatrans.pr,'wetfreq'),is=i)))
grid()
```

@@

```{r}
eof.es <- EOF(es)
eof.slp <- EOF(slp)
index(eof.es) <- year(eof.es)
index(eof.slp) <- year(eof.slp)
```

@@

@@

@@

## **Examine the residuals from the downscaling**

The residuals for the wet-day frequency:

@@

#### **Do the downscaling based on the CMIP5 RCP4.5 ensemble**

@@

```{r}
for (i in 1:3) {
  qqnorm(coredata(subset(subset(tx.sd,it='jja'),is=i)),main=loc(tx.sd)[i])
}
```

>>>>

```{r,echo=TRUE}
library(esd)
library(ncdf)
#rm(list=ls())
## Information about the system and session
```

@@

### **Daily station data**
```{r,echo=TRUE}
## Load precipitation data extracted from the gridded IMD data set:
load('climatrans.pr.rda')
## Display the structure of the data as
str(climatrans.pr)
plot(climatrans.pr,new=FALSE)
```

##**Sensitivity tests**

@@

```{r}
wq95 <- function(x) {x <- x[is.finite(x)]; x <- x[x >= 1]; wq95 <- quantile(x,probs=0.95); wq95}
season<-5:10
fw <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetfreq')
mu <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wetmean')
nd <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'nv')
ndhr <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'count',threshold=100)
q95 <- aggregate(subset(climatrans.pr,it=month.abb[season]),year,'wq95')
```

@@

```{r, fig.width=6, fig.height=6}
## Probability of exceedin 50 mm/day
Pr <- PrexpPr(mu,fw,x0=50) 
plot(Pr,new=FALSE,errorbar=FALSE)
for(i in 1:3) lines(trend(subset(Pr,is=i)))
grid()
print(trend(Pr,result='pval'))
```

@@

```{r, fig.width=6, fig.height=6}
## Aggregate to annual totals
pt <- annual(climatrans.pr,'sum')
attr(pt,'unit') <- 'mm/season'
plot(pt,new=FALSE)
for(i in 1:3) lines(trend(subset(pt,is=i)))
grid()
print(trend(pt,result='pval'))
```

Also, it's interesting to see how the number of days with heave rainfall changes over time:

```{r, fig.width=6, fig.height=6}
plot(ndhr,new=FALSE)
grid()
```

@@

```{r}
eof.es <- EOF(es)
eof.slp <- EOF(slp)
eof.t2m <- EOF(aggregate(subset(t2m,is=list(lon=c(65,80),lat=c(10,20))),year,'mean'))
index(eof.es) <- year(eof.es)
index(eof.slp) <- year(eof.slp)
index(eof.t2m) <- year(eof.t2m)
```

@@

The main May-October T(2m) mode was associated with the strength of the Monsoon winds.

### **Plot the EOFs for the SLP**
```{r, fig.width=6, fig.height=6}
plot(eof.t2m,new=FALSE)
```

@@

### Examine the residuals from the downscaling for the wet-day mean

### Alternative approach: downscale the number of days with heavy rainfall

Model the number with heavy rainfall (more than 100 mm/day) in Mumbay.

```{r fig.width=6, fig.height=6}
ndhr.cal <- data.frame(y = coredata(subset(subset(ndhr,is=2),it=c(1949,2004))),
                       X = coredata(subset(eof.es,it=c(1949,2004))))
ds.ndhr <- glm(y ~ X.1 + X.2 + X.3 + X.4 + X.5 + X.6 + X.7, data=ndhr.cal,family="poisson")
plot(1949:2004,ndhr.cal$y,pch=19,ylab=expression(N(X > 100*mm)),xlab='',main=loc(ndhr)[2])
grid()
lines(1949:2004,exp(predict(ds.ndhr)),col='red')
```

Check if the number of days with heavy precipitation is Poisson-distributed

```{r}
print(paste('Mean number=',mean(ndhr[,2],na.rm=TRUE),'Variance=',var(ndhr[,2],na.rm=TRUE),
            'should equal',mean(ndhr[,2],na.rm=TRUE),'for Poisson process'))
h <- hist(coredata(ndhr[,2]),freq=FALSE,col='grey',
     main=loc(ndhr)[2])
lines(h$mids-0.5,dpois(x=h$mids-0.5,lambda=mean(ndhr[,2],na.rm=TRUE)),lwd=2,col='red')
grid()
```

@@

### The residuals for the wet-day frequency:

@@

## **Do the downscaling based on the CMIP5 RCP4.5 ensemble**

@@

```{r}
for (i in 1:3) {
  qqnorm(coredata(subset(subset(tx.sd,it='jja'),is=i)),main=loc(tx.sd)[i])
  qqline(coredata(subset(subset(tx.sd,it='jja'),is=i)),main=loc(tx.sd)[i])
}
```
